### 2019-06-07 paper submission 3tex.tex
Make minor edits after proofreading.

\title{Diversification through Trade \thanks{%
Thanks: Referees, Pol Antras, Costas Arkolakis, Robert Barro, Fernando
Broner, Ariel Burstein, Lorenzo Caliendo, Julian DiGiovanni, Bernardo
Guimaraes, Nobu Kiyotaki, Pete Klenow, David Laibson, Fabrizio Perri, Steve
Redding, Ina Simonovska, Jaume Ventura, Romain Wacziarg, and seminar
participants at Bocconi, Birmingham, CREI, Princeton, Penn, Yale, NYU, UCL,
LBS, LSE, Toulouse, Warwick, as well as participants at SED, ESSIM, and the
Nottingham trade conference. Calin Vlad Demian, Balazs Kertesz, Federico
Rossi, and Peter Zsohar provided superb research assistance. Caselli
acknowledges financial support from the Luverlhume Fellowship. Koren
acknowledges financial support from the European Research Council (ERC)
starting grant 313164. Tenreyro acknowledges financial support from the ERC
starting grant 240852. \newline
$^{\dag }$ London School of Economics, CfM, CEPR. $^{\ddag }$ Central
European University, MTA KRTK, CEPR. $^{\S }$ European Comission.
Correspondence: s.tenreyro@LSE.ac.uk.} }
\author{Francesco Caselli$^{\dag }$ \and Miklos Koren$^{\ddag }$ \and Milan
Lisicky$^{\S }$ \and Silvana Tenreyro$^{\dag }$}
\date{This draft: November 2017}
\maketitle

@@

We use the model in conjunction with sector-level production and bilateral
trade data for a diverse group of 24 countries to quantitatively assess how
changes in trading costs since the early 1970s have affected income
volatility.\footnote{%
The data are disaggregated into 24 sectors. We stop the analysis in 2007 as
our model abstracts from the factors underlying the financial crisis.} We
find that the decline in trade costs since the 1970s has caused sizeable
reductions in income volatility in the vast majority of\textbf{\ }the
countries in our sample. The range of changes in volatility due to trade
varies significantly across countries, with the largest reductions being
nearly 80\%. On average, volatility fell 37\% compared to a counterfactual
where trade barriers remain at their early-1970s level.

@@

The focus of our quantitative evaluation is real income, defined as nominal
GDP deflated by a cost of living index. In the model, the cost of living
index is a preferenced-based ideal price deflator. In the data counterpart,
the cost of living index is the CPI. Hence, we work with a welfare-relevant
notion of income.\footnote{%
Kehoe and Ruhl (2008) and Burstein and Cravino (2015) study the theoretical
impact of foreign productivity shocks on various measures of domesitc
economic activity. In general, foreign productivity shocks (or other sources
of change in the terms of trade) have little first-order effects on
production-based measures of activity (e.g. GDP deflated by the GDP
deflator), while they have first-order effects on welfare-based measures.}
Indeed, we could alternatively have focused on consumption volatility,
because, for most countries in the world, income and consumption
fluctuations are almost perfectly correlated, and in our model income equals
consumption. For the same reason\textbf{, }our model abstracts from trade in
financial assets.

@@

A paper that is closely related to ours is Burgess and Donaldson (2012), who
use the Eaton-Kortum model in conjunction with data on the expansion of
railroads across regions in India to assess whether real income became more
or less sensitive to rainfall shocks, as India's regions became more open to
trade. The authors find that the decline in transportation costs lowered the
impact of productivity shocks on real income, implying a reduction in
volatility. Our analysis is at a higher level of generality,\textbf{\ }and
highlights that, while a reduction in volatility has been experienced by
many countries as they became more open to trade, the size and sign of the
trade effect on volatility may be--- and indeed has been---different across
different countries.\footnote{%
See also Donaldson (2015), where the question is also addressed in the
context of India's railroad expansion. There is\ also a growing literature
on the effect of globalization on income risk and inequality. We do not
focus on distributional effects within countries in this paper, though it is
obviously a very important issue, and a natural next step in our research.
For theoretical developments in that area, see for example, Anderson (2011)
and the references therein.}

\textbf{While this paper focuses on contrasting our new
diversification-through-trade mechanism with the traditional sectoral
specialization mechanism, it leaves to future work to include the role of
\textquotedblleft granular\textquotedblright\ shocks. As pointed out by di
Giovanni and Levchenko (2012), if openness to trade increases concentration
the impact of granular shocks is exacerbated, potentially leading to an
increase in volatility. See also di Giovanni, Levchenko and Mejean (2014)
for a country-level application.}

@@

At a given point in time $t$, country $n$ is endowed with $L_{nt}$ units of
a primary (non produced) input, which we interpret as equipped labour. At
the beginning of each period, before the realization of the shocks $%
A_{nt}^{j}$, a representative consumer decides on the optimal allocation of
the primary input $L_{nt}$ across the different sectors, $L_{nt}^{j}$. After
the shocks to productivity are realized, equipped labour can be reallocated
within a sector, but not across sectors. Next, production and consumption
take place. Clearly, clearing in the input market within a sector implies 
\begin{equation*}
L_{nt}^{j}=\int_{0}^{1}l_{nt}(\omega ^{j})d\omega ^{j}.
\end{equation*}

@@

Using (\ref{BC}) in the utility function we can solve for the sectoral labor
allocation:%
\begin{equation}
L_{nt}^{j}=\arg \max E_{t-1}\left[ \log \left( \frac{%
\sum_{j=1}^{J}w_{nt}^{j}L_{nt}^{j}}{P_{nt}}\right) \right]
,s.t.:\sum\nolimits_{j=1}^{J}L_{nt}^{j}=L_{nt},  \label{eq:log:utility}
\end{equation}%
where $E_{t-1}$ indicates the rational expectation over
the possible realizations of period $t$ shocks. In particular, the
representative agent knows the previous values of the shock processes $%
A_{nt-1}^{j}$. The agent also knows the distribution of $A_{nt}^{j}$%
 conditional on $A_{nt-1}^{j}$ (which we estimate later
from data), and is therefore able to compute the rational expectation in (%
\ref{eq:log:utility}).\footnote{%
Implicit in our formulation is the assumption that there is perfect
risk-sharing within a country, but no risk-sharing across countries. To
motivate the lack of risk-sharing across countries, see our earlier
discussion on the high comovement between consumption and output as well as
the high correlation between consumption and output volatility.}

@@

\begin{equation}
P_{nt}=\prod\nolimits_{j}^{J}\left( \frac{1}{\alpha _{n}^{j}}\right)
^{\alpha ^{j}}\left( P_{nt}^{j}\right) ^{\alpha ^{j}},  \label{eq1}
\end{equation}%
\begin{equation}
R_{nt}^{j}=\sum\nolimits_{m=1}^{N}d_{mnt}^{j}E_{mt}^{j},  \label{rev}
\end{equation}%
\begin{equation}
E_{nt}^{j}=\alpha _{t}^{j}P_{nt}C_{nt}+\sum\nolimits_{k=1}^{J}\gamma
^{jk}R_{nt}^{k},  \label{exp}
\end{equation}%
\begin{equation}
w_{nt}^{j}L_{nt}^{j}=\beta ^{j}R_{nt}^{j},  \label{RC}
\end{equation}%
and the budget constraint (\ref{BC}). In the equations above, $d_{nmt}^{j}$
is the fraction of country $n$'s total spending on sector-$j$ goods that is
imported from country $m;$ $P_{nt}^{j}$ is the price of sectoral-good $j$ in
country $n$; $R_{nt}^{j}$ is total revenues accruing to firms operating in
sector $j$ in country $n$; and $E_{nt}^{j}$ is total expenditure by country $%
n$ residents (consumers and firms) on sectoral good $j$. $B^{j}\equiv \left(
\beta ^{j}\right) ^{-\beta ^{j}}\dprod\limits_{k=1}^{J}\left( \gamma
^{k_{j}}\right) ^{-\gamma ^{k_{j}}}$ and $\xi \equiv \left[ \Gamma \left( 
\frac{\theta +1-\eta }{\theta }\right) \right] $, where $\Gamma $ is the
gamma function, are parametric constants. Hence, equation (\ref{shares})
says that country $n$ imports disproportionately from countries $m$ and
sectors $j$ that have high productivity draws $T_{m}^{j}$ and $A_{mt}^{j}$;
low wages $w_{mt}^{j}$ and sectoral prices $P_{mt}^{k}$; and low bilateral
trading costs, namely high $\kappa _{nmt}$s. Equation (\ref{eq2}) says that
the same factors affect domestic sectoral prices. Equation (\ref{eq1})
follows from the final-good producer's profit maximization problem, and
shows the price of consumption as an aggregate of the sectoral prices.
Equation (\ref{rev}) expresses the total sales of sector $j$ in country $n$
as a function of each country's expenditures on that sector and the share of
country $n$ in each country's imports in that sector. Equation (\ref{exp})
states that a country's expenditures in sector $j$ is the sum of final and
intermediate uses of sector $j$ goods. Equation (\ref{RC}) simply notes from
the Cobb-Douglas formulation that value added from sector $j$ is a share $%
\beta ^{j}$ of the gross output of sector $j$.

@@

Under complete autarky, it can be easily shown that value added in the
one-sector economy is a function of augmented productivity: 
\begin{equation*}
Y_{nt}\propto \left( Z_{nt}\right) ^{\frac{1}{\beta \theta }}
\end{equation*}%
where, recall, $Z_{nt}\equiv T_{n}\left( L_{nt}A_{nt}^{1/\beta }\right)
^{\beta \theta }$. Defining $\hat{Z}_{nt}$ ($\hat{Y}_{nt}$) as the
log-deviation of $Z_{nt}$ ($Y_{nt}$) from its deterministic trend, we thus
have $\hat{Y}_{nt}=\frac{1}{\beta \theta }\hat{Z}_{nt}.$ Hence, much as in a
RBC model, in the one-sector economy under autarky shocks to value added are
driven exclusively by domestic shocks to the productive capacity of the
economy, $\hat{Z}_{nt}.$ The variance of income, $Var(\hat{Y}_{nt})$ thus
depends on the variance of the shocks $Var(\hat{Z}_{nt})$: 
\begin{equation*}
Var(\hat{Y}_{nt})=\frac{1}{\left( \beta \theta \right) ^{2}}Var(\hat{Z}%
_{nt}).
\end{equation*}

@@

Figure 2 starts by comparing the baseline model-generated income volatility
with the volatility in the data. The baseline model uses our benchmark
calibration, $\theta =4$, and feeds in the historical time series for the
trade costs $\kappa _{mnt}$, and for the augmented productivity factors $%
Z_{nt}^{j}$. The graph shows the standard deviation of real income
deviations from trend. Recall that real income is measured as value added
deflated by the expenditure-based price index. The data counterpart is
nominal GDP deflated by the CPI index. The correlation between standard
deviations in the model and data series is 0.96 (0.88 without China) for the
standard deviation and 0.99 (0.89 without China) for the variance. The
analysis that follows will focus on the variance as a measure of volatility,
rather than the standard deviation, because we exploit the additivity
properties of the former to separately account for the diversification and
sectoral-specialization effects.

@@

The comparison in Column 1 reveals that volatility is generally higher under
the counterfactual scenario with constant trading costs than in the
baseline. For all countries except for China, there would have been more
volatility under constant trade costs than there has been. For almost all
countries, therefore, the common wisdom which predicts greater volatility
following trade integration does not seem to apply.

The biggest declines in volatility caused by trade occurred in
Belgium-Luxemburg, Canada, Denmark, Germany, Ireland, Mexico, the
Netherlands, Spain, and the United Kingdom, all of which saw volatility
reductions due to trade in excess of 50\% (meaning their volatility has been
50 percent lower than it would have been had trading costs stayed at their
1972 levels). In the one country in which trade has created additional
volatility, the excess volatility is negligible. The (unweighted) average
country in our sample experienced a 36\% decline in volatility thanks to
increased openness. But this average effect masks a huge amount of
heterogeneity in the quantitative and qualitative effect of trade in
volatility, consistent with our discussion of the country-specificity of the
trade-volatility relation.

@@

Because of the additive properties of the variance, the specialization
effect can be quantified as the difference between the overall change in
volatility, and the change due to the diversification effect. This is
reported in Column 3. The figures should be interpreted as the change in
volatility due to trade integration that would have occurred if only
sectoral shocks (global or country specific) had been present. The change is
positive for 15 out of 24 countries. Consistent with the standard view,
therefore, the specialization channel tends to increase volatility in a
majority of cases. Remarkably, however, there is a large number of countries
which evidently are pushed to specialize into less volatile sectors, or into
sectors that comove negatively (or less positively) with the country's
aggregate shocks or other sectoral shocks. On average, the specialization
channel implies an increase in volatility of 5\%.

The most important lesson from the comparison of Columns 2 and 3 is about
the relative magnitude of the diversification and specialization effects.
The average change due to the diversification mechanism is about eight times
as large as the average change due to the specialization mechanism. The
specialization effect, on which the policy debate seems centred, is not as
important as the diversification effect. We have hinted at the likely reason
for this in the Introduction: country-specific shocks are simply much more
important quantitatively than sector-specific ones.

@@

\textbf{This overall monotonic decline in volatilty, however, masks some
more nuanced dynamics of the diversification and specialization effects. In
particular, the diversification effect peters out in the period 2000-2007.
This petering out in the last seven years of the sample may possibly reflect
some noisiness due to the relative short time span over which volatilities
are computed. However, taken at face value, it points to the fact that --
consistent with our theory -- the impact of trade on volatility is not only
heterogenous across countries, but also over time. For example, the decline
in the diversification effect could be due to country-wide shocks becaming
more correlated in the 2000s.}

@@

In this section we evaluate the robustness of our baseline results to three
alternative implementation choices: (i) allowing for unbalanced trade; (ii)
alternative calibration values; (iii) allowing for costly labor reallocation
across sectors, \textbf{and (iv) allowing for elasticities of substitution
in consumption other than 1.}

@@

In this section we use our model to investigate two further questions about
the forces at work in our model and in the data. In particular we ask: (i)
What is the quantitative role of intersectoral input-output linkages in the
relationship between trade openness and volatility? And (ii) Did the
emergence of China as a global trading powerhouse exert a disproportionate
effect in other countries' volatility through trade?

@@

To see if input-output linkages do indeed amplify the impact of trade on
income volatility in our model, we compare our baseline results to those of
an alternative model without intermediates, i.e. where we set $\gamma
^{kj}=0 $ for all $j$ and $k$ (and consequently $\beta =1$). We then
re-calibrate the productivity shocks and the trade cost processes to fit
value-added and trade data, as before. The results from this no-input-output
model are presented in Table $7$, and should as usual be compared to those
of Table 1. While the qualitative findings are similar to those of the full
model with input-output linkages, the quantitative impact of trade is
considerably reduced in their absence. The average decline in volatility due
to trade is only 3.5\% (as usual entirely by the diversification effect).
Hence, allowing firms to source inputs from other sectors is crucial to
capture the full effects of trade on volatility.

@@

\subsubsection*{Inner Loop}
Introduce an auxiliary variable for the factory-gate price of intermediate goods,
\begin{equation}\label{eq:varrho}
	\phi_{nt}^j = \xi B^j (T_{n}^j)^{-1/\theta} (A_{nt}^j)^{-1} w_{nt}^{j\beta^j} \prod_{k=1}^J P_{nt}^{k\gamma^{kj}}.
\end{equation}
The prices of the final goods and wage rates can be directly expressed as a function of intermediate prices and parameters,
\begin{equation}\label{eq:price}
	P_{mt}^j(\phi) = \left[
		\sum_{n=1}^N (\phi_{nt}^j/\kappa_{mnt}^j)^{-\theta}
		\right]^{-1/\theta}.
\end{equation}
\begin{equation}\label{eq:wage}
	w_{nt}^j(\phi) = (\xi B^j)^{-1/\beta^j}
	(\phi_{nt}^j)^{1/\beta^j}  
	(T_{n}^j)^{1/(\beta^j\theta)}(A_{nt}^j)^{1/\beta^j} 
	\prod_{k=1}^J P_{nt}^k(\phi)^{-\gamma^{kj}/\beta^j}.
\end{equation}
Sectoral revenues are a constant multiple of sectoral wage bill, so, given labor allocations, we can also express them as a function of only intermediate prices and parameters,
\begin{equation}\label{eq:revenue}
	R_{nt}^j(\phi) = w_{nt}^j(\phi)L_{nt}^j/\beta^j.
\end{equation}
Taking the market clearing conditions in \eqref{rev} and \eqref{exp}, we can write intermediate prices as a function of sectoral revenues,
\begin{equation}\label{eq:market_clearing}
	\phi_{nt}^j = R_{nt}^j(\phi)^{-1/\theta}
	\left\{
	\sum_{m=1}^N
		[\kappa_{mnt}^j P_{mt}^j(\phi)]^{\theta}
		\sum_{k=1}^J
			\left( \alpha_{t}^j
				\beta^k 
			+ \gamma^{jk}\right)R_{mt}^k(\phi)
	\right\}^{1/\theta}.
\end{equation}
We start from a guess for sectoral revenues (computed analytically for the free-trade equilibrium) and then use \eqref{eq:market_clearing} to obtain a guess for intermediate prices. We recompute sectoral revenues under these intermediate prices and iterate until convergence.

>>>>

\title{Diversification through Trade \thanks{%
Thanks: Referees, Pol Antras, Costas Arkolakis, Robert Barro, Fernando
Broner, Ariel Burstein, Lorenzo Caliendo, Julian DiGiovanni, Bernardo
Guimaraes, Nobu Kiyotaki, Pete Klenow, David Laibson, Fabrizio Perri, Steve
Redding, Ina Simonovska, Jaume Ventura, Romain Wacziarg, and seminar
participants at Bocconi, Birmingham, CREI, Princeton, Penn, Yale, NYU, UCL,
LBS, LSE, Toulouse, Warwick, as well as participants at SED, ESSIM, and the
Nottingham trade conference. Calin Vlad Demian, Balazs Kertesz, Federico
Rossi, and Peter Zsohar provided superb research assistance. Caselli
acknowledges financial support from the Luverlhume Fellowship. Koren
acknowledges financial support from the European Research Council (ERC)
starting grant 313164. Tenreyro acknowledges financial support from the ERC
starting grant 240852. \newline
$^{\dag }$ London School of Economics, CfM, CEPR. $^{\ddag }$ Central
European University, MTA KRTK, CEPR. $^{\S }$ European Comission.
Correspondence: s.tenreyro@LSE.ac.uk.} }
\author{Francesco Caselli$^{\dag }$ \and Mikl\'os Koren$^{\ddag }$ \and Milan
Lisicky$^{\S }$ \and Silvana Tenreyro$^{\dag }$}
\date{This draft: June 2019}
\maketitle

@@

We use the model in conjunction with sector-level production and bilateral
trade data for a diverse group of 24 countries to quantitatively assess how
changes in trading costs since the early 1970s have affected income
volatility.\footnote{%
The data are disaggregated into 24 sectors. We stop the analysis in 2007 as
our model abstracts from the factors underlying the financial crisis.} We
find that the decline in trade costs since the 1970s has caused sizeable
reductions in income volatility in the vast majority of the
countries in our sample. The range of changes in volatility due to trade
varies significantly across countries, with the largest reductions being
nearly 80\%. On average, volatility fell 36\% compared to a counterfactual
where trade barriers remain at their early-1970s level.

@@

The focus of our quantitative evaluation is real income, defined as nominal
GDP deflated by a cost of living index. In the model, the cost of living
index is a preferenced-based ideal price deflator. In the data counterpart,
the cost of living index is the CPI. Hence, we work with a welfare-relevant
notion of income.\footnote{%
Kehoe and Ruhl (2008) and Burstein and Cravino (2015) study the theoretical
impact of foreign productivity shocks on various measures of domesitc
economic activity. In general, foreign productivity shocks (or other sources
of change in the terms of trade) have little first-order effects on
production-based measures of activity (e.g. GDP deflated by the GDP
deflator), while they have first-order effects on welfare-based measures.}
Indeed, we could alternatively have focused on consumption volatility,
because, for most countries in the world, income and consumption
fluctuations are almost perfectly correlated, and in our model income equals
consumption. For the same reason, our model abstracts from trade in
financial assets.

@@

A paper that is closely related to ours is Burgess and Donaldson (2012), who
use the Eaton-Kortum model in conjunction with data on the expansion of
railroads across regions in India to assess whether real income became more
or less sensitive to rainfall shocks, as India's regions became more open to
trade. The authors find that the decline in transportation costs lowered the
impact of productivity shocks on real income, implying a reduction in
volatility. Our analysis is at a higher level of generality, and
highlights that, while a reduction in volatility has been experienced by
many countries as they became more open to trade, the size and sign of the
trade effect on volatility may be--- and indeed has been---different across
different countries.\footnote{%
See also Donaldson (2015), where the question is also addressed in the
context of India's railroad expansion. There is\ also a growing literature
on the effect of globalization on income risk and inequality. We do not
focus on distributional effects within countries in this paper, though it is
obviously a very important issue, and a natural next step in our research.
For theoretical developments in that area, see for example, Anderson (2011)
and the references therein.}

While this paper focuses on contrasting our new
diversification-through-trade mechanism with the traditional sectoral
specialization mechanism, it leaves to future work to include the role of
\textquotedblleft granular\textquotedblright\ shocks. As pointed out by di
Giovanni and Levchenko (2012), if openness to trade increases concentration
the impact of granular shocks is exacerbated, potentially leading to an
increase in volatility. See also di Giovanni, Levchenko and Mejean (2014)
for a country-level application.

@@

At a given point in time $t$, country $n$ is endowed with $L_{nt}$ units of
a primary (non produced) input, which we interpret as equipped labour. At
the beginning of each period, before the realization of the shocks $%
A_{nt}^{j}$, a representative consumer decides on the optimal allocation of
the primary input $L_{nt}$ across the different sectors, $L_{nt}^{j}$. After
the shocks to productivity are realized, equipped labour can be reallocated
within a sector, but not across sectors. Next, production and consumption
take place. Clearing in the input market within a sector implies 
\begin{equation*}
L_{nt}^{j}=\int_{0}^{1}l_{nt}(\omega ^{j})d\omega ^{j}.
\end{equation*}

@@

Using (\ref{BC}) in the utility function we can solve for the sectoral labor
allocation:%
\begin{equation}
L_{nt}^{j}=\arg \max E_{t-1}\left[ \log \left( \frac{%
\sum_{j=1}^{J}w_{nt}^{j}L_{nt}^{j}}{P_{nt}}\right) \right]
,s.t.:\sum\nolimits_{j=1}^{J}L_{nt}^{j}=L_{nt},  \label{eq:log:utility}
\end{equation}%
where $E_{t-1}$ indicates the rational expectation over
the possible realizations of period $t$ shocks. In particular, the
representative agent knows the previous values of the shock processes $%
A_{nt-1}^{j}$. The agent also knows the distribution of $A_{nt}^{j}$ 
conditional on $A_{nt-1}^{j}$ (which we estimate later
from data), and is therefore able to compute the rational expectation in (%
\ref{eq:log:utility}).\footnote{%
Implicit in our formulation is the assumption that there is perfect
risk-sharing within a country, but no risk-sharing across countries. To
motivate the lack of risk-sharing across countries, see our earlier
discussion on the high comovement between consumption and output as well as
the high correlation between consumption and output volatility.}

@@

\begin{equation}
P_{nt}=\prod\nolimits_{j}^{J}\left( \frac{1}{\alpha _{n}^{j}}\right)
^{\alpha ^{j}}\left( P_{nt}^{j}\right) ^{\alpha ^{j}},  \label{eq1}
\end{equation}%
\begin{equation}
R_{nt}^{j}=\sum\nolimits_{m=1}^{N}d_{mnt}^{j}E_{mt}^{j},  \label{rev}
\end{equation}%
\begin{equation}
E_{nt}^{j}=\alpha _{t}^{j}P_{nt}C_{nt}+\sum\nolimits_{k=1}^{J}\gamma
^{jk}R_{nt}^{k},  \label{exp}
\end{equation}%
\begin{equation}
w_{nt}^{j}L_{nt}^{j}=\beta ^{j}R_{nt}^{j},  \label{RC}
\end{equation}%
and the budget constraint (\ref{BC}). In the equations above, $d_{nmt}^{j}$
is the fraction of country $n$'s total spending on sector-$j$ goods that is
imported from country $m;$ $P_{nt}^{j}$ is the price of sectoral-good $j$ in
country $n$; $R_{nt}^{j}$ is total revenues accruing to firms operating in
sector $j$ in country $n$; and $E_{nt}^{j}$ is total expenditure by country $%
n$ residents (consumers and firms) on sectoral good $j$. $B^{j}\equiv \left(
\beta ^{j}\right) ^{-\beta ^{j}}\dprod\limits_{k=1}^{J}\left( \gamma
^{k_{j}}\right) ^{-\gamma ^{k_{j}}}$ and $\xi \equiv \Gamma \left( 
\frac{\theta +1-\eta }{\theta }\right)$, where $\Gamma $ is the
gamma function, are parametric constants. Hence, equation (\ref{shares})
says that country $n$ imports disproportionately from countries $m$ and
sectors $j$ that have high productivity draws $T_{m}^{j}$ and $A_{mt}^{j}$;
low wages $w_{mt}^{j}$ and sectoral prices $P_{mt}^{k}$; and low bilateral
trading costs, namely high $\kappa _{nmt}$s. Equation (\ref{eq2}) says that
the same factors affect domestic sectoral prices. Equation (\ref{eq1})
follows from the final-good producer's profit maximization problem, and
shows the price of consumption as an aggregate of the sectoral prices.
Equation (\ref{rev}) expresses the total sales of sector $j$ in country $n$
as a function of each country's expenditures on that sector and the share of
country $n$ in each country's imports in that sector. Equation (\ref{exp})
states that a country's expenditures in sector $j$ is the sum of final and
intermediate uses of sector $j$ goods. Equation (\ref{RC}) simply notes from
the Cobb-Douglas formulation that value added from sector $j$ is a share $%
\beta ^{j}$ of the gross output of sector $j$.

@@

Under complete autarky, it can be easily shown that value added in the
one-sector economy is a function of augmented productivity: 
\begin{equation*}
Y_{nt}\propto \left( Z_{nt}\right) ^{\frac{1}{\beta \theta }}
\end{equation*}%
where, recall, $Z_{nt}\equiv T_{n}\left( L_{nt}A_{nt}^{1/\beta }\right)
^{\beta \theta }$. Defining $\hat{Z}_{nt}$ ($\hat{Y}_{nt}$) as the
log-deviation of $Z_{nt}$ ($Y_{nt}$) from its deterministic trend, we thus
have $\hat{Y}_{nt}=\frac{1}{\beta \theta }\hat{Z}_{nt}.$ Hence, much as in an
RBC model, in the one-sector economy under autarky shocks to value added are
driven exclusively by domestic shocks to the productive capacity of the
economy, $\hat{Z}_{nt}.$ The variance of income, $Var(\hat{Y}_{nt})$ thus
depends on the variance of the shocks $Var(\hat{Z}_{nt})$: 
\begin{equation*}
Var(\hat{Y}_{nt})=\frac{1}{\left( \beta \theta \right) ^{2}}Var(\hat{Z}%
_{nt}).
\end{equation*}

@@

Figure 2 starts by comparing the baseline model-generated income volatility
with the volatility in the data. The baseline model uses our benchmark
calibration, $\theta =4$, and feeds in the historical time series for the
trade costs $\kappa _{mnt}$, and for the augmented productivity factors $%
Z_{nt}^{j}$. The graph shows the standard deviation of real income
deviations from trend. Recall that real income is measured as value added
deflated by the expenditure-based price index. The data counterpart is
nominal GDP deflated by the CPI index. The correlation between volatility in the model and data series is 0.96 (0.88 without China) for the
standard deviation and 0.99 (0.89 without China) for the variance. The
analysis that follows will focus on the variance as a measure of volatility,
rather than the standard deviation, because we exploit the additivity
properties of the former to separately account for the diversification and
sectoral-specialization effects.

@@

The comparison in Column 1 reveals that volatility is generally higher under
the counterfactual scenario with constant trading costs than in the
baseline. For all countries except for China and the Rest of the World, there would have been more
volatility under constant trade costs than there actually was. For almost all
countries, therefore, the common wisdom which predicts greater volatility
following trade integration does not seem to apply.

The biggest declines in volatility caused by trade occurred in
Belgium-Luxemburg, Canada, Denmark, Germany, Ireland, Mexico, the
Netherlands, Spain, and the United Kingdom, all of which saw volatility
reductions due to trade in excess of 50\% (meaning their volatility has been
50 percent lower than it would have been had trading costs stayed at their
1972 levels). In the two countries-regions in which trade has created additional
volatility, the excess volatility is negligible. The (unweighted) average
country in our sample experienced a 36\% decline in volatility thanks to
increased openness. But this average effect masks a huge amount of
heterogeneity in the quantitative and qualitative effect of trade in
volatility, consistent with our discussion of the country-specificity of the
trade-volatility relation.

@@

Because of the additive properties of the variance, the specialization
effect can be quantified as the difference between the overall change in
volatility, and the change due to the diversification effect. This is
reported in Column 3. The figures should be interpreted as the additional change in
volatility due to trade integration that would have occurred with the presence of
sectoral shocks (global or country specific). The change is
positive for 13 out of 25 countries. Consistent with the standard view,
therefore, the specialization channel tends to increase volatility in a
majority of cases. Remarkably, however, there is a large number of countries
which evidently are pushed to specialize into less volatile sectors, or into
sectors that comove negatively (or less positively) with the country's
aggregate shocks or other sectoral shocks. On average, the specialization
channel implies an increase in volatility of 5\%.

The most important lesson from the comparison of Columns 2 and 3 is about
the relative magnitude of the diversification and specialization effects.
The average change due to the diversification mechanism is about eight times
as large, in absolute value, as the average change due to the specialization mechanism. The
specialization effect, on which the policy debate seems centred, is not as
important as the diversification effect. We have hinted at the likely reason
for this in the Introduction: country-specific shocks are simply much more
important quantitatively than sector-specific ones.

@@

This overall monotonic decline in volatilty, however, masks some
more nuanced dynamics of the diversification and specialization effects. In
particular, the diversification effect peters out in the period 2000-2007.
This petering out in the last seven years of the sample may possibly reflect
some noisiness due to the relative short time span over which volatilities
are computed. However, taken at face value, it points to the fact that --
consistent with our theory -- the impact of trade on volatility is not only
heterogenous across countries, but also over time. For example, the decline
in the diversification effect could be due to country-wide shocks becaming
more correlated in the 2000s.

@@

In this section we evaluate the robustness of our baseline results to three
alternative implementation choices: (i) allowing for unbalanced trade; (ii)
alternative calibration values; (iii) allowing for costly labor reallocation
across sectors, and (iv) allowing for elasticities of substitution
in consumption other than 1.

@@

In this section we use our model to investigate two further questions about
the forces at work in our model and in the data. In particular we ask: (i)
What is the quantitative role of intersectoral input-output linkages in the
relationship between trade openness and volatility? And (ii) Did the
emergence of China as a global trading powerhouse exert a disproportionate
effect on other countries' volatility through trade?

@@

To see if input-output linkages do indeed amplify the impact of trade on
income volatility in our model, we compare our baseline results to those of
an alternative model without intermediates, i.e. where we set $\gamma
^{kj}=0 $ for all $j$ and $k$ (and consequently $\beta^j =1$). We then
re-calibrate the productivity shocks to fit
value-added and trade data, as before. The results from this no-input-output
model are presented in Table $7$, and should as usual be compared to those
of Table 1. While the qualitative findings are similar to those of the full
model with input-output linkages, the quantitative impact of trade is
considerably reduced in their absence. The average decline in volatility due
to trade is only 3.5\% (as usual entirely by the diversification effect).
Hence, allowing firms to source inputs from other sectors is crucial to
capture the full effects of trade on volatility.

@@

\subsubsection*{Inner Loop}
Introduce an auxiliary variable for the factory-gate price of intermediate goods,
\begin{equation}\label{eq:varrho}
	\phi_{nt}^j = \xi B^j (T_{n}^j)^{-1/\theta} (A_{nt}^j)^{-1} w_{nt}^{j\beta^j} \prod_{k=1}^J P_{nt}^{k\gamma^{kj}}.
\end{equation}
The prices of the final goods and wage rates can be directly expressed as a function of intermediate prices and parameters,
\begin{equation}\label{eq:price}
	P_{mt}^j(\phi) = \left[
		\sum_{n=1}^N (\phi_{nt}^j/\kappa_{mnt}^j)^{-\theta}
		\right]^{-1/\theta}.
\end{equation}
\begin{equation}\label{eq:wage}
	w_{nt}^j(\phi) = (\xi B^j)^{-1/\beta^j}
	(\phi_{nt}^j)^{1/\beta^j}  
	(T_{n}^j)^{1/(\beta^j\theta)}(A_{nt}^j)^{1/\beta^j} 
	\prod_{k=1}^J P_{nt}^k(\phi)^{-\gamma^{kj}/\beta^j}.
\end{equation}
Sectoral revenues are a constant multiple of sectoral wage bill, so, given labor allocations, we can also express them as a function of only intermediate prices and parameters,
\begin{equation}\label{eq:revenue}
	R_{nt}^j(\phi) = w_{nt}^j(\phi)L_{nt}^j/\beta^j.
\end{equation}
Taking the market clearing conditions in \eqref{rev} and \eqref{exp}, we can write intermediate prices as a function of sectoral revenues,
\begin{equation}\label{eq:market_clearing}
	\phi_{nt}^j = R_{nt}^j(\phi)^{-1/\theta}
	\left\{
	\sum_{m=1}^N
		[\kappa_{mnt}^j P_{mt}^j(\phi)]^{\theta}
		\sum_{k=1}^J
			\left( \alpha_{t}^j
				\beta^k 
			+ \gamma^{jk}\right)R_{mt}^k(\phi)
	\right\}^{1/\theta}.
\end{equation}
We start from a guess for sectoral revenues (computed analytically for the free-trade equilibrium) and then use \eqref{eq:market_clearing} to obtain a guess for intermediate prices. We recompute sectoral revenues under these intermediate prices and iterate until convergence. Given a solution for intermediate prices, we have a closed form solution for wage rates and other prices.
